As is usually the case, this thesis would not have been possible without a lot of people. I would like to thank as many as possible here, but my apologies in advance if I miss anyone. First and foremost, I would like to thank Stephan for giving me the opportunity to work in the awesome Grill lab. I'll be forever grateful for both his scientific guidance and feedback throughout the project and his guidance and encouragement for my personal development. I would like to thank him especially for supporting me through the tough times during the start of my PhD. I would also like to thank my TAC committee, Frank J{\"u}licher and Elisabeth Fischer-Friedrich, for their criticisms and advice on the project, which helped in its development. I also thank the LMF (especially Britta) for amazing microscope support, Carolyn from the international office for support in dealing with housing and visa, and Ivan Baines for his help during the earlier parts of my PhD.

A big thanks goes to Peter Gro{\ss}, who started this project before I joined with Mirna Kramar and Michael Nestler, and introduced me to the wonderful world of experimental biophysics with \ac{ce} worms. Peter patiently taught me how to work with \ac{ce} worms, how to use (and not break) the microscope and subsequent analysis of the images taken during microscopy. I am very thankful for Peter's guidance in both scientific and non-scientific matters. I also thank Mark Leaver for his guidance through the later parts of the PhD project, for his help on genetic manipulations in \ac{ce}, and for patiently listening to my ramblings. Thanks also to my collaborators from TU Dresden, Michael Nestler and Axel Voigt, for a fruitful collaboration and the numerous numerical simulations that support and advance the work described in this thesis. I would also like to thank all those who corrected and provided feedback on my thesis, in spite of their busy schedule.

The Grill lab is an amazing place to work in, thanks in no small part to the amazing people that make up the lab. Thanks to everyone in the lab for making this awesome place -- Peter, Lokesh, Arjun, Teije, Argo, Mark, Saurabh, J{\'u}lia, Victoria, J{\'o}se, Sandeep, Christoph, Pranay, Tina Wiegand, Julia Pfanzelter, Rana, Yahor, Andrea, Kyle, Debayan, Jamie, Adrian, Patrick, Arittri, Jan, Isabel, Lutz, Gisela -- I apologize if I missed anyone. I thank Lokesh for his \enquote{life advices}, and for the board game nights. Thanks to Arjun for all the discussions, and for letting me ramble on various topics during our walks back from work to home. I also thank the Dutch faux-Jesus of the lab, Teije, who was always ready to help with his genetics knowledge and more. Thanks to Victoria and Argo for helpful discussions on PhD life and beyond. Thanks to Andrea, Debayan and Jamie for help with the microscope after the GPU surgery, and also to Jamie for his advice on writing. I would also like to thank Yahor and Arjun for the evergreen discussions on crazy experimental and theory ideas. Thanks also to the technicians -- Karin, Tina Neumann and Friedricke -- who helped me settle in the lab at the start, and were always ready to help me out with any experimental setup, however crazy it sounded. The Grill lab was (and is) my second home away from home, and I would thank everyone in the lab for making it so. 

I would also like to thank the Brian, Birgit and Cassandra from the PhD office, who helped me navigate the intricacies of the bureaucracy of the PhD program and that of the TU Dresden. Thanks also the IMPRS 2018 selection batch, who became some of my closest friends in Dresden -- Adrian, Isabel, Aryaman, Manan, Abhijeet and others. I will never forget the crazy and wonderful time I had during the PhD party preparation with them. I would also like to thank Kaushik, Anatol, Huang, Matheus and Joy for their help and advice at the start of my PhD that led me to the Grill lab -- and thanks especially to Kaushik without whom I would have struggled in my early days in Dresden. I would also like to thank the MPI-CBG board game group of Regis, Albine, Tyler, Ilker and others that introduced me to board games. Thanks also to Pandemic board games group of Ilker, Lokesh and Tyler, who kept me sane during the pandemic in 2020 and after. I would also like to acknowledge \enquote{Elbe ke albele}: Aryaman, Manan, Abhijeet and Alishan, who have been amazing friends during my stay here in Dresden.

Finally, I would like to thank my parents and sister, without whose support and encouragement I would never be here in the first place -- and to whom this thesis is dedicated. I do not have the words to express my thanks to them, and am deeply indebted to their support.